---
title: "Political Consequences of The Prison Industrial Compex"
author: "Mya Zepp"
format: revealjs
theme: simple
editor: visual
css: style.css
---

## Disenfranchisement Laws in The United States

```{r}
library(usmap)
library(leaflet)
library(here)
library(tidyverse)
library(ggplot2)
library(broom)
library(dplyr)
library(usmap)
library(modelsummary)
library(poliscidata)

disenfran_laws <- read_csv(
  here::here("disenfranchisement_laws_data.csv")
)

colnames(disenfran_laws) <- c("state", "value")

disenfran_laws$state <- tolower(disenfran_laws$state)


disenfran_laws$value <- as.factor(disenfran_laws$value)

plot_usmap(data = disenfran_laws, values = "value") +
  scale_fill_manual(
    values = c("orange", "red", "darkred", "purple", "darkblue", "blue"),
    labels = c(
      "No disenfranchisement",
      "Voting rights restored upon release",
      "Voting rights restored for those on probation or parole 
who have not been incarcerated in the last 5 years",
      "Voting rights restored after completion of sentence 
(including probation or parole)",
      "Permanent disenfranchisement for some",
      "Permanent disenfranchisement for all"
    )
  ) +
  theme_minimal() +  
  theme(
    legend.position = "right",
    panel.grid = element_blank(), 
    panel.background = element_blank(), 
    axis.ticks = element_blank(),  
    axis.text = element_blank()  
  ) +
  labs(
    title = "Disenfranchisement Laws in the United States",
    caption = "Source: Brennan Center for Justice, 2024"
  )
```

# To what extent does prison time impact an individual's feelings about politics and their desire to vote?

## Hypothesis

$$H_0$$ Prison time has no impact on an individuals feelings about politics and their desire to vote

$$H_1$$ Prison time negatively impacts an individuals feelings about politics and their desire to vote

## Data

::: incremental
-   2020 Marshall Project and Slate Survey of 8,000 incarcerated individuals across Arkansas, California, Illinois, Kansas, Montana, Maine

-   Respondents:

    -   3,873 White
    -   974 Black
    -   435 Hispanic

-   The survey, being the first of its kind has many limitations
:::

## Trust in Political Officials

```{r}
prison_data <- read_csv(
  here::here("incarcerated_survey_2019.csv")
) |>
  drop_na()

prison_data <- prison_data |>
  mutate(
    how_often_officials_acting_in_your_interest = case_when(
      how_often_officials_acting_in_your_interest %in% c("Always", "Very Often") ~ 1,
      how_often_officials_acting_in_your_interest %in% c("Never", "Rarely") ~ 0,
      TRUE ~ NA_real_
    )
  ) |>
  drop_na(how_often_officials_acting_in_your_interest, length_in_this_facility) 

ggplot(prison_data, aes(x = how_often_officials_acting_in_your_interest)) +
  geom_bar(fill = "blue") +
  labs(
    title = "Trust in Political Officials",
    x = "How often are officials acting in your interest",
    y = "Count"
  ) +
   theme_minimal() +
  theme(legend.position = "top")


```

## Regression Equation

$$
Motivation to Vote = \alpha + \beta_1T +\beta_2*Length in Facility + \beta_3 * Identifies As White + \beta_4 * Identifies As Hispanic or Latino + \beta_5 * Identifies As Black + \beta_6* Age +\epsilon
$$

## Results

```{r}
prison_data <- read_csv(
  here::here("incarcerated_survey_2019.csv")
) |>
  drop_na()
  
prison_cleaned <- prison_data |>
  filter(!is.na(length_in_this_facility)) |>
  filter(!is.na(incarceration_impacts_motivation_to_vote))


prison_cleaned$length_in_this_facility <- ifelse(prison_cleaned$length_in_this_facility == "10 years or less", 1, 0)
#1 = ten years or less

prison_cleaned$age <- ifelse(prison_cleaned$age == "35 or younger", 1, 0) 

prison_cleaned$identifies_as_black <-  ifelse(prison_cleaned$identifies_as_black == "TRUE", 1, 0)

prison_cleaned$identifies_as_white <-  ifelse(prison_cleaned$identifies_as_white == "TRUE", 1, 0)

prison_cleaned$identifies_as_hispanic_or_latinx <-  ifelse(prison_cleaned$identifies_as_hispanic_or_latinx == "TRUE", 1, 0)

prison_cleaned <- prison_cleaned %>%
  mutate(incarceration_impacts_motivation_to_vote = case_when(
    incarceration_impacts_motivation_to_vote %in% c("Increased my motivation to vote", "Slightly increased my motivation to vote") ~ 1, incarceration_impacts_motivation_to_vote %in% c("No impact", "Decreased my motivation to vote",  "Slightly decreased my motivation to vote") ~ 0
  )) 

mod <- lm(incarceration_impacts_motivation_to_vote ~ length_in_this_facility + identifies_as_white + identifies_as_hispanic_or_latinx + identifies_as_black + age, data = prison_cleaned)

#tidy(mod)

modelsummary(mod,
             title = "Impact of Incarceration on Voting Motivation",
             coef_map = c(
               "length_in_this_facility" = "Length in Facility",
               "identifies_as_white" = "White",
               "identifies_as_hispanic_or_latinx" = "Hispanic/Latinx",
               "identifies_as_black" = "Black",
               "age" = "Age"
             ),
             stars = TRUE,
             statistic = "std.error",
          
)

##need to add figures on a plot with predicted points
```

## Conclusion

-   More time spent in prison is associated with a decreased motivation to vote
