---
title: "Political Consequences of the Prison Industrial Complex"
author: "Mya Zepp"
format: revealjs
theme: simple
editor: visual
---

## Definitions

-   Disenfranchisement

-   Prison Industrial Complex

## Disenfranchisement Laws in The United States

```{r}
library(usmap)
library(leaflet)
library(here)
library(tidyverse)
library(ggplot2)
library(broom)
library(dplyr)
library(usmap)
library(modelsummary)
library(poliscidata)
library(ggpattern)

disenfran_laws <- read_csv(
  here::here("disenfranchisement_laws_data.csv")
)

colnames(disenfran_laws) <- c("state", "value")

disenfran_laws$state <- tolower(disenfran_laws$state)


disenfran_laws$value <- as.factor(disenfran_laws$value)

plot_usmap(data = disenfran_laws, values = "value") +
  scale_fill_manual(
    values = c("orange", "red", "darkred", "purple", "darkblue", "blue"),
    labels = c(
      "No disenfranchisement",
      "Voting rights restored upon release",
      "Voting rights restored for those on probation or parole 
who have not been incarcerated in the last 5 years",
      "Voting rights restored after completion of sentence 
(including probation or parole)",
      "Permanent disenfranchisement for some",
      "Permanent disenfranchisement for all"
    )
  ) +
  theme_minimal() +  
  theme(
    legend.position = "right",
    panel.grid = element_blank(), 
    panel.background = element_blank(), 
    axis.ticks = element_blank(),  
    axis.text = element_blank()  
  ) +
  labs(
    title = "Disenfranchisement Laws in the United States",
    caption = "Source: Brennan Center for Justice, 2024"
  )
```

## Existing Literature

-   “If there are more than 4 million people who are legally disenfranchised, there may be an equal number who believe they are ineligible to vote, thus depressing registration and voting participation.” (Drucker and Barreras, 2005, p. 10)
-   “Custodial interactions negatively affect the likelihood of participating in politics and carrying out the responsibilities of citizenship.” (Weaver and Lerman, 2010, p. 818)
-   "The system of confinement is highly effective at keeping people separated, at breaking the possibility of meaningful relationships” (Pearson, 2024, p.56)

# To what extent does prison time impact an individual's feelings about politics and their desire to vote?

## Hypothesis

$$H_0$$ Prison time has no impact on an individuals feelings about politics and their desire to vote

$$H_1$$ Prison time negatively impacts an individuals feelings about politics and their desire to vote

## Data

-   2020 Marshall Project and Slate Survey of 8,000 incarcerated individuals across Arkansas, California, Illinois, Kansas, Montana, Maine (most respondents are in Kansas or Arkansas)

    -   2157 usable responses

-   Respondents:

    -   1,859 White
    -   272 Black
    -   26 Hispanic

-   The survey, being the first of its kind has many limitations

## Results

```{r}
prison_data <- read_csv(
  here::here("incarcerated_survey_2019.csv")
) |>
  drop_na()
  
prison_cleaned <- prison_data |>
  filter(!is.na(length_in_this_facility)) |>
  filter(!is.na(incarceration_impacts_motivation_to_vote)) 


prison_cleaned$length_in_this_facility <- ifelse(prison_cleaned$length_in_this_facility == "10 years or less", 0, 1)
#more time = 1

prison_cleaned$age <- ifelse(prison_cleaned$age == "36 or older", 1, 0)
#old = 1

prison_cleaned$identifies_as_black <-  ifelse(prison_cleaned$identifies_as_black == "TRUE", 1, 0)

prison_cleaned$identifies_as_white <-  ifelse(prison_cleaned$identifies_as_white == "TRUE", 1, 0)

prison_cleaned <- prison_cleaned |>
  mutate(
    how_often_officials_acting_in_your_interest = case_when(
      how_often_officials_acting_in_your_interest %in% c("Always", "Very Often") ~ 1,
      how_often_officials_acting_in_your_interest %in% c("Never", "Rarely") ~ 0,
      TRUE ~ NA_real_
    )
  )

prison_cleaned$identifies_as_hispanic_or_latinx <-  ifelse(prison_cleaned$identifies_as_hispanic_or_latinx == "TRUE", 1, 0)

prison_cleaned$ever_voted <-  ifelse(prison_cleaned$ever_voted == "Yes", 1, 0)

prison_cleaned <- prison_cleaned %>%
  mutate(incarceration_impacts_motivation_to_vote = case_when(
    incarceration_impacts_motivation_to_vote %in% c("Increased my motivation to vote", "Slightly increased my motivation to vote") ~ 1, incarceration_impacts_motivation_to_vote %in% c("No impact", "Decreased my motivation to vote",  "Slightly decreased my motivation to vote") ~ 0
  )) 

prison_cleaned <- prison_cleaned %>%
  filter(party %in% c("The Democratic party", "The Republican party", "Independent")) %>%
  mutate(party = case_when(
    party == "Independent" ~ 2,
    party == "The Democratic party" ~ 1,
    party == "The Republican party" ~ 0
  ))

unique(prison_cleaned$party)

mod <- lm(incarceration_impacts_motivation_to_vote ~ length_in_this_facility + ever_voted + identifies_as_hispanic_or_latinx + identifies_as_black + age + party, data = prison_cleaned)

##mod_1 <- lm(how_often_officials_acting_in_your_interest ~ length_in_this_facility + ever_voted + identifies_as_hispanic_or_latinx + identifies_as_black + age, data = prison_cleaned)

##summary(mod_1)


#tidy(mod)

modelsummary(mod,
             title = "Impact of Incarceration on Voting Motivation",
             coef_map = c(
               "length_in_this_facility" = "Length in Facility",
               "ever_voted" = "Voted in the past",
               #"identifies_as_hispanic_or_latinx" = "Hispanic/Latinx",
              # "identifies_as_black" = "Black",
              # "age" = "Age",
               "how_often_officials_acting_in_your_interest" = "Politicians are acting in your interest",
               "party" = "Party"
             ),
             stars = TRUE,
             gof_map = c("nobs", "r.squared")
          
)
```

## Discussion of Results

## Conclusion

## Sources 
