---
title: "Phase 2 Data Analysis - Mya Zepp"
format: html
editor: visual
---

## **GVPT 729A: Phase 2 Guidelines for Preliminary Results Assignment**

## 1) A clear definition of the dependent variable with details on how it is measured and its source.

In 2020, the Marshall Project and Slate partnered to gather information on the politics of people behind bars. They conducted a voluntary survey of 8,000 incarcerated individuals across six states. The majority of respondents were white men in Kansas and Arkansas and thus not representative of the overall prison population. Thus, any conclusions drawn from the data are only truly generalizable to white incarcerated people in Republican states. The methods used in conducting this survey are another area of interest. The surveyors had no direct contact with prisons or imprisoned people; rather, they conducted their research through a tablet company that provides tablets to incarcerated people. The survey was given to this company and then sent to the tablets assigned to individual people. The Marshall Project and Slate are journalistic in nature, so they had no reason to go through IRB or conduct research directly with the prisons, but it does raise questions as to if this data would be different if they had. That being said, this data is the first of its kind and opens doors for future explorations.

Using this data, I explore the answer provided to the question "What impact has incarceration had on your motivation to vote?" to which respondents answered 'increased my motivation to vote', 'slightly increased my motivation to vote', 'decreased my motivation to vote', 'slightly decreased my motivation to vote,' or 'no impact.' For the purposes of this analysis, I excluded 'no impact' (n = 996). I compare the results of this question to the length of time the respondents have spent incarcerated, which is coded as '10 years or less' and 'more than 10 years'

## 2) A clear statement the main hypothesis(es) - if there are more than two discuss this with me before handing anything in. Included here should be a clear definition of the independent variables mentioned in the hypotheses and an indication of how they are measured.

$H_0$: Spending 10 years or more in prison will negatively impact or not impact motivation to vote

$H_1$: Spending 10 years or more in prison will positively impact motivation to vote

## 3) Discussion of the relationship between the primary independent variable(s) and the dependent variable.  Here, based on the nature of the measurement of the dependent variable and the independent variables I expect you to explain what the best way to get a handle on this relationship is in the bivariate case and to display in graphical and/or tabular form the results (professionally formatted as always). There might be several ways that are all pretty good; if that is the case say so, explaining why there isn’t one single best approach, and discuss just one or two approaches (depending on how much space you have) and place the other graphs or tables in an appendix (as always, these too should be formatted professionally as if they were to appear in an academic journal). You should also make sure that the variables of primary interest will not cause problems with respect to model identification. If this can be seen from the results you present be sure to note whether the evidence suggests problems or not; if the presentation method you selected as best does not naturally reveal information of this sort, perform the appropriate test, report the results in a footnote and include the relevant table(s) or graph(s) in the appendix.

## 4) An appendix with a table of descriptive statistics for the dependent variable and the primary independent variables, formatted as usual - just present the table with the sample size, mean, standard deviation, minimum, and maximum.

```{r}
#| echo: false

library(usmap)
library(leaflet)
library(here)
library(tidyverse)
library(ggplot2)
library(broom)
library(usmap)
library(modelsummary)
library(ggpattern)


prison_data <- read_csv(
  here("incarcerated_survey_2019.csv"),
  show_col_types = FALSE
) |>
  drop_na()

prison_cleaned <- prison_data |>
  filter(!is.na(length_in_this_facility)) |>
  filter(!is.na(incarceration_impacts_motivation_to_vote)) 
table(prison_cleaned$incarceration_impacts_motivation_to_vote)
table(prison_cleaned$length_in_this_facility)

```

## 5) Do the write-up (i.e. numbers 1-3 above) in no more than 2 double spaced pages with 12 point font, and 1 inch margins - you can have a separate page with your name etc. to save space for the write-up. Be sure to answer the following key question: Is there support for your hypothesis(es) based on your bivariate analysis? Put the tables and/or graphs that you discuss on separate pages (these do not count against the 2 page limit for text).

## Appendix

```{r}
#| echo: false


library(usmap)
library(leaflet)
library(here)
library(tidyverse)
library(ggplot2)
library(broom)
#library(dplyr)
library(usmap)
library(modelsummary)
library(ggpattern)


prison_data <- read_csv(
  here("incarcerated_survey_2019.csv"),
  show_col_types = FALSE
) |>
  drop_na()

prison_cleaned <- prison_data |>
  filter(!is.na(length_in_this_facility)) |>
  filter(!is.na(incarceration_impacts_motivation_to_vote)) 



prison_cleaned$length_in_this_facility <- ifelse(prison_cleaned$length_in_this_facility == "10 years or less", 0, 1)
#more time = 1



prison_cleaned <- prison_cleaned %>%
  mutate(incarceration_impacts_motivation_to_vote = case_when(
    incarceration_impacts_motivation_to_vote %in% c("Increased my motivation to vote", "Slightly increased my motivation to vote") ~ 1, incarceration_impacts_motivation_to_vote %in% c("Decreased my motivation to vote",  "Slightly decreased my motivation to vote") ~ 0
  )) 

prison_cleaned <- prison_cleaned %>%
  mutate(incarceration_motivation_label = case_when(
    incarceration_impacts_motivation_to_vote == 1 ~ "Increased motivation",
    incarceration_impacts_motivation_to_vote == 0 ~ "Decreased motivation"
  )) |>
  mutate(
    length_label = case_when(
      length_in_this_facility == 1 ~ "More than 10 years",
      length_in_this_facility == 0 ~ "10 years or less"
    )
  ) |>
  rename(
    `Time incarcerated` = length_label,
    `Motivation to vote` = incarceration_motivation_label
  )


datasummary_crosstab(`Time incarcerated` ~ `Motivation to vote`, 
                     data = prison_cleaned,)
                     

mod <- lm(incarceration_impacts_motivation_to_vote ~ length_in_this_facility, data = prison_cleaned)

modelsummary(mod,
             title = "Impact of Time Incarcerated on Voting Motivation",
             coef_map = c(
               "length_in_this_facility" = "Length in Facility"
           ),
             stars = TRUE,
             gof_map = c("nobs", "r.squared")
           
          
)
```
