---
title: "GVPT729A Phase 3 - Mya Zepp"
format: pdf: 
    fontsize: 12pt
    linestretch: 2 
    mainfont: "Times New Roman"
    margin-bottom: 1in
    margin-top: 1in
    margin-left: 1in
    margin-right: 1in
    keep-tex: true
    include-in-header:
      text: |
        % Required by modelsummary/tabularray
        \usepackage{tabularray}
        \usepackage{float}
        \usepackage{graphicx}
        \usepackage{codehigh}
        \usepackage[normalem]{ulem}
        \UseTblrLibrary{booktabs}
        \UseTblrLibrary{siunitx}
editor: visual
---

\begin{center}

GVPT 729A: Phase 2 Preliminary Results Assignment

Mya Zepp

October 27, 2025

\end{center}

\pagebreak

## 1) Run your model(s) (in as close a form to the final models as possible).

```{r}
#| echo: false
#Slide 1

library(foreign)
library(usmap)
library(leaflet)
suppressPackageStartupMessages(library(tidyverse))
library(ggplot2)
library(broom)
library(usmap)
library(modelsummary)
library(ggpattern)
library(mvtnorm)  


prison_data <- read_csv(
  "C:/Users/mzepp/Downloads/School/R/Research/incarcerated_survey_2019.csv",
  show_col_types = FALSE
) |>
  drop_na()


prison_cleaned <- prison_data |>
  filter(!is.na(length_in_this_facility)) |>
  filter(!is.na(incarceration_impacts_motivation_to_vote)) 



prison_cleaned$length_in_this_facility <- ifelse(prison_cleaned$length_in_this_facility == "10 years or less", 0, 1)
#more time = 1



prison_cleaned <- prison_cleaned %>%
  mutate(incarceration_impacts_motivation_to_vote = case_when(
    incarceration_impacts_motivation_to_vote %in% c("Increased my motivation to vote", "Slightly increased my motivation to vote") ~ 1, incarceration_impacts_motivation_to_vote %in% c("Decreased my motivation to vote",  "Slightly decreased my motivation to vote", "no impact") ~ 0
  )) 

prison_cleaned$age <- ifelse(prison_cleaned$age == "36 or older", 1, 0)
#old = 1

prison_cleaned$identifies_as_black <-  ifelse(prison_cleaned$identifies_as_black == "TRUE", 1, 0)

prison_cleaned$identifies_as_white <-  ifelse(prison_cleaned$identifies_as_white == "TRUE", 1, 0)

prison_cleaned <- prison_cleaned |>
  mutate(
    how_often_officials_acting_in_your_interest = case_when(
      how_often_officials_acting_in_your_interest %in% c("Always", "Very Often") ~ 1,
      how_often_officials_acting_in_your_interest %in% c("Never", "Rarely") ~ 0,
      TRUE ~ NA_real_
    )
  )

prison_cleaned$identifies_as_hispanic_or_latinx <-  ifelse(prison_cleaned$identifies_as_hispanic_or_latinx == "TRUE", 1, 0)

prison_cleaned$ever_voted <-  ifelse(prison_cleaned$ever_voted == "Yes", 1, 0)

prison_cleaned <- prison_cleaned %>%
  filter(party %in% c("The Democratic party", "The Republican party", "Independent")) %>%
  mutate(party = case_when(
    party == "Independent" ~ 1,
    party == "The Democratic party" ~ 2,
    party == "The Republican party" ~ 0
  ))

prison_cleaned <- prison_cleaned %>%
  mutate(incarceration_motivation_label = case_when(
    incarceration_impacts_motivation_to_vote == 1 ~ "Increased motivation",
    incarceration_impacts_motivation_to_vote == 0 ~ "Decreased motivation/No change"
  )) |>
  mutate(
    length_label = case_when(
      length_in_this_facility == 1 ~ "More than 10 years",
      length_in_this_facility == 0 ~ "10 years or less"
    )
  ) |> 
     rename(
    `Time incarcerated` = length_label,
    `Motivation to vote` = incarceration_motivation_label
  )

#probit.model <- glm(`Time incarcerated` ~ `Motivation to vote`, family=binomial(link="probit"), data=prison_cleaned)

##probit.model <- glm(incarceration_impacts_motivation_to_vote ~ length_in_this_facility + identifies_as_black  + identifies_as_hispanic_or_latinx + ever_voted + party + age*length_in_this_facility, family = binomial(link = "probit"),data = prison_cleaned)

probit.model <- glm(incarceration_impacts_motivation_to_vote ~ length_in_this_facility + identifies_as_black  + identifies_as_hispanic_or_latinx + ever_voted + party, 
                    family = binomial(link = "probit"),
                    data = prison_cleaned)

#summary(probit.model)


modelsummary(probit.model,
             title = "Impact of Incarceration on Voting Motivation",
             coef_map = c(
               "length_in_this_facility" = "Length in Facility",
               "ever_voted" = "Voted in the past",
               "identifies_as_black" = "Black",
               "identifies_as_hispanic_or_latinx" = "Hispanic/Latino",
               "age" = "Age",
               "how_often_officials_acting_in_your_interest" = "Politicians are acting in your interest",
               "party" = "Party"
             ),
             stars = TRUE,
             gof_map = c("nobs", "r.squared")
          
)


# y = motivation 
# x = time 

preds <- predict(probit.model, type = "response")
summary(preds)


coefs_probit <- coef(probit.model)
coefs_probit

pprobit <- pnorm(coefs_probit[2]*prison_cleaned$length_in_this_facility + coefs_probit[3]*prison_cleaned$identifies_as_black + coefs_probit[4]*prison_cleaned$identifies_as_hispanic_or_latinx + coefs_probit[5]*prison_cleaned$ever_voted + coefs_probit[6]*prison_cleaned$party + coefs_probit[1])
summary(pprobit)

testp <- preds - pprobit
summary(testp)

px1_1 <- pnorm(coefs_probit[2]*1 + coefs_probit[3]*prison_cleaned$identifies_as_black + coefs_probit[4]*prison_cleaned$identifies_as_hispanic_or_latinx + coefs_probit[5]*prison_cleaned$ever_voted + coefs_probit[6]*prison_cleaned$party + coefs_probit[1])
summary(px1_1)


px1_0 <- pnorm(coefs_probit[2]*0 + coefs_probit[3]*prison_cleaned$identifies_as_black + coefs_probit[4]*prison_cleaned$identifies_as_hispanic_or_latinx + coefs_probit[5]*prison_cleaned$ever_voted + coefs_probit[6]*prison_cleaned$party + coefs_probit[1])
summary(px1_0)


effectx1 <- px1_1 - px1_0


summary(cbind(px1_1, px1_0, effectx1))

# the effect of length in facility is 1.2 percentage points

```

```{r}
#| echo: false

#Simulation

m1 <- glm(incarceration_impacts_motivation_to_vote ~ length_in_this_facility + identifies_as_black  + identifies_as_hispanic_or_latinx + ever_voted + party, 
                    family = binomial(link = "probit"),
                    data = prison_cleaned)
summary(m1)

coefs <- coef(m1)
coefs


d <- prison_cleaned[complete.cases(prison_cleaned),]


n_draws <- 1000

set.seed(17)
sim_coefs <- rmvnorm(n_draws, coefs, vcov(m1)) 

dim(sim_coefs)
colnames(sim_coefs)

rbind(coefs, apply(sim_coefs, 2, mean)) 

p_mean <- NULL
for (i in 1:1000) {
  p_mean[i] <- mean(pnorm(sim_coefs[i,1] + sim_coefs[i,2]*d$length_in_this_facility + sim_coefs[i,3]*d$identifies_as_black + sim_coefs[i,4]*d$identifies_as_hispanic_or_latinx + sim_coefs[i,5]*d$ever_voted + sim_coefs[i,6]*d$party))
}

summary(p_mean)

px1_1_mean <- NULL
for (i in 1:1000) {
  px1_1_mean[i] <- mean(pnorm(sim_coefs[i,1] + sim_coefs[i,2]*1 + sim_coefs[i,3]*d$identifies_as_black + sim_coefs[i,4]*d$identifies_as_hispanic_or_latinx + sim_coefs[i,5]*d$ever_voted + sim_coefs[i,6]*d$party))
}


px1_0_mean <- NULL
for (i in 1:1000) {
  px1_0_mean[i] <- mean(pnorm(sim_coefs[i,1] + sim_coefs[i,2]*0 + sim_coefs[i,3]*d$identifies_as_black + sim_coefs[i,4]*d$identifies_as_hispanic_or_latinx + sim_coefs[i,5]*d$ever_voted + sim_coefs[i,6]*d$party))
}

effectx1_mean <- px1_1_mean - px1_0_mean
summary(effectx1_mean)

margeffx2_mean <- NULL
for (i in 1:1000) {
  margeffx2_mean[i] <- mean(dnorm(sim_coefs[i,1] + sim_coefs[i,2]*d$length_in_this_facility + sim_coefs[i,3]*d$identifies_as_black)*sim_coefs[i,3] *sim_coefs[i,4] *sim_coefs[i,5]*sim_coefs[i,6])
}

lapply(means <- list(p_mean, px1_1_mean, px1_0_mean, effectx1_mean, margeffx2_mean), summary)


q <- do.call("rbind", (lapply(means, quantile, c(.025,.975)))) 
q                                                              

results.p <- cbind(q[,1], lapply(means, mean),q[,2]) 
colnames(results.p) <- c("2.5", "Mean", "97.5") 
rownames(results.p) <- c("p_mean", "px1_1_mean", "px1_0_mean", "effectx1_mean", "margeffx2_mean") 
results.p
```

## 2) Produce the relevant predictions from at least one model including p values or confidence intervals around at least one of the predictions related to your hypotheses.

## 3) Provide files with the code and data you used to accomplish steps 1 & 2 without any extra code for recodes, failed attempts etc. 

## 4) Write a report that states the main hypothesis (if there is more than 1, just pick 1 hypothesis), discusses the nature of the dependent variable, discusses the model used, and discusses the support for your hypothesis.  The report should be no more than 2 double spaced pages with 12 point font, and 1 inch margins - you can have a separate page with your name etc. to save space for the write-up.  Put the tables and/or graphs that you discuss on separate pages (these do not count against the 2 page limit for text). substantive significance

You can add options to executable code like this

The `echo: false` option disables the printing of code (only output is displayed).
