---
title: ""
format: 
  pdf:
    fontsize: 12pt
    linestretch: 2 
    mainfont: "Times New Roman"
    margin-bottom: 1in
    margin-top: 1in
    margin-left: 1in
    margin-right: 1in
    keep-tex: true
    include-in-header:
      text: |
        % Required by modelsummary/tabularray
        \usepackage{tabularray}
        \usepackage{float}
        \usepackage{graphicx}
        \usepackage{codehigh}
        \usepackage[normalem]{ulem}
        \UseTblrLibrary{booktabs}
        \UseTblrLibrary{siunitx}
editor: visual
---

\begin{center}

GVPT 729A: Phase 2 Preliminary Results Assignment

Mya Zepp

October 27, 2025

\end{center}

\pagebreak

In 2020, the Marshall Project and Slate partnered to gather information on the politics of people behind bars. They conducted a voluntary survey of 8,000 incarcerated individuals across six states. The majority of respondents were white men in Kansas and Arkansas, and thus not representative of the overall prison population. As such, any conclusions drawn from the data are only truly generalizable to white incarcerated men in Republican states. The methods used in conducting this survey are another area of interest. The surveyors had no direct contact with prisons or imprisoned people; rather, they conducted their research through a tablet company that provides tablets to incarcerated people. The survey was given to this company and then sent to the tablets assigned to individual people.

Using this data, I explore the answer provided to the question "What impact has incarceration had on your motivation to vote?" to which respondents answered, 'increased my motivation to vote', 'slightly increased my motivation to vote', 'decreased my motivation to vote', 'slightly decreased my motivation to vote,' or 'no impact.' For the purposes of this analysis, I excluded 'no impact' (n = 996). I compare the results of this question to the length of time the respondents have spent incarcerated, which is coded as '10 years or less' and 'more than 10 years'

Previous research in the field of incarceration and politics has had mixed opinions on the impact of the prison industrial complex on voting. One study indicated that any interaction with the carceral state, even a traffic ticket, can lead to decreased voter turnout (Weaver and Lerman, 2010) while another suggests that incarceration causes no real change in civic participation (Gerber et al., 2017) and a minority of research argues, as I will, that some people may in fact be mobilized by interactions with the carceral state (Walker, 2020). The data I plan to explore examines the impact of incarceration on 'motivation,' not participation. As such, we cannot make inferences on how these same respondents will participate, but I hypothesize, similar to Hannah Walker, that spending more than 10 years in prison will positively impact motivation to vote ($H_1$). In the analysis that follows I examine these variables through a cross tabulation and through an OLS regression model. In future analysis I plan to use binary logit to further explore the relationship and I plan to include controls for race, party identification, age, and having voted in the past.

Preliminary analysis of the variables demonstrates a relationship that supports my hypothesis. The cross-tabulation of responses shows that among respondents who have spent 10 years or less incarcerated, 45.8% stated their motivation to vote increased, and respondents who have spent more than 10 years incarcerated had similar results, with 58.8% stating their motivation to vote increased. Length in facility demonstrates that spending 10 years or more in prison is associated with an increase in motivation to vote compared to those who have spent less than 10 years in prison. This result is in line with the argument that “a pivotal factor connecting experiences with the criminal justice system to political mobilization is a sense of systemic injustice.” (Walker, 2020, p. 6) This association could be a result of more time spent considering the implications of the prison-industrial complex, resulting in a stronger desire to participate politically. It is also possible that losing the right to vote makes one have a stronger desire to vote once they get that right back.

The data suggests that motivation to vote may be high among incarcerated people, demonstrating that the conventional wisdom suggesting that previously incarcerated people are not participating in politics due to a lack of information may very well be true because based on this data, it is not due to a lack of motivation.

## Appendix

$H_0$: Spending 10 years or more in prison will negatively impact or not impact motivation to vote

$H_1$: Spending 10 years or more in prison will positively impact motivation to vote

```{r}
#| echo: false


library(usmap)
library(leaflet)
suppressPackageStartupMessages(library(tidyverse))
library(ggplot2)
library(broom)
library(usmap)
library(modelsummary)
library(ggpattern)



#suppressPackageStartupMessages(library(conflicted))
#conflicts_prefer(dplyr::filter)

prison_data <- read_csv(
  "C:/Users/mzepp/Downloads/School/R/Research/incarcerated_survey_2019.csv",
  show_col_types = FALSE
) |>
  drop_na()



prison_cleaned <- prison_data |>
  filter(!is.na(length_in_this_facility)) |>
  filter(!is.na(incarceration_impacts_motivation_to_vote)) 



prison_cleaned$length_in_this_facility <- ifelse(prison_cleaned$length_in_this_facility == "10 years or less", 0, 1)
#more time = 1



prison_cleaned <- prison_cleaned %>%
  mutate(incarceration_impacts_motivation_to_vote = case_when(
    incarceration_impacts_motivation_to_vote %in% c("Increased my motivation to vote", "Slightly increased my motivation to vote") ~ 1, incarceration_impacts_motivation_to_vote %in% c("Decreased my motivation to vote",  "Slightly decreased my motivation to vote", "no impact") ~ 0
  )) 

prison_cleaned <- prison_cleaned %>%
  mutate(incarceration_motivation_label = case_when(
    incarceration_impacts_motivation_to_vote == 1 ~ "Increased motivation",
    incarceration_impacts_motivation_to_vote == 0 ~ "Decreased motivation/No change"
  )) |>
  mutate(
    length_label = case_when(
      length_in_this_facility == 1 ~ "More than 10 years",
      length_in_this_facility == 0 ~ "10 years or less"
    )
  ) |> 
     rename(
    `Time incarcerated` = length_label,
    `Motivation to vote` = incarceration_motivation_label
  )


```

#### Summary: Incarceration Impact on Motivation to Vote

```{r}
#| echo: false

summary(prison_cleaned$incarceration_impacts_motivation_to_vote)

```

#### Summary: Time Spent in Prison

```{r}
#| echo: false
#| 

summary(prison_cleaned$length_in_this_facility)


mod <- lm(incarceration_impacts_motivation_to_vote ~ length_in_this_facility, data = prison_cleaned)

options("modelsummary_format_numeric_latex" = "plain")

modelsummary(mod,
             title = "Time Incarcerated Regressed on Voting Motivation",
             coef_map = c(
               "length_in_this_facility" = "Length in Facility"
           ),
             stars = TRUE,
             gof_map = c("nobs", "r.squared")
           
          
)



datasummary_crosstab(`Time incarcerated` ~ `Motivation to vote`, data = prison_cleaned, title ="Time Incarcerated and Voting Motivation", output = "latex")

```
